### 2009-04-14 doc/storage_system/arc-storage-documentation.tex
Went through all of Alex' comments

The different replicas of a file could be in different \note[JKN]{TODO: which states? a reference to definition of states maybe?}states, e.g. the replica could be valid and alive or just in the process of creation or it could be corrupt or a whole storage node could be offline. This state is always stored as metadata of the given replica. For each file there is a checksum calculated, and this checksum is used to detect if a replica gets corrupted. If a storage node (more precisely: the Shepherd service on the storage node) detects that a file is invalid, it reports this so the metadata will be in sync with the real state. And the storage nodes sends heartbeat messages periodically, and if a storage node goes offline, the missing heartbeat triggers the modification of metadata as well.

@@

\begin{enumerate}
    \item The client tool needs to know the URL of a Bartender. This could be preconfigured on the client side or in future releases it could be acquired from an information system. When the client tool has the URL, it sends a `list' request which contains the Logical Name \verb!/ourvo/users/user!.
    \item The Bartender tries to get the metadata of the given LN by sending a `traverseLN' request to a Librarian.
    \item The Librarian service starts the traversing by asking an A-Hash service about the first part of the LN, which is the \verb!/! root collection. The A-Hash service only knows about GUID's and not about LN's, but the GUID of the root collection is well-known, so the A-Hash can return the metadata of the list of files and sub-collections in the root collection.
    \item Hopefully, the \verb!ourvo! collection can be found in the root collection, which means that the Librarian knows its GUID, and can ask for its metadata from the A-Hash.
    \item After the A-Hash returns the metadata of the \verb!/ourvo! collection, the Librarian finds the GUID of \verb!users! in it, then gets its metadata.
    \item The A-Hash returns the metadata of \verb!/ourvo/users! which contains the GUID of \verb!user!, so the Librarian can ask for its metadata.
    \item At last the A-Hash returns the metadata of \verb!/ourvo/users/user! to the Librarian, and the Librarian returns it to the Bartender. This metadata contains the list of entries within this collection, and it also contains the access policies for this collection.
    \item The Bartender first checks, based on the user's DN (or his VO membership) and the access policies of this collection, if the user has permission to get the contents of this collection or not. If the user is approved, then because the `list' request should return additional metadata about each entry in the collection, the Bartender sends a `get' message to a Librarian which requests the metadata of all the entries in this collection.
    \item The Librarian gets the metadata from the A-Hash and returns it to the Bartender.
    \item Now the Bartender has all the needed information and returns the proper response to the client tool which formats and prints the results nicely for the user.
\end{enumerate}

@@

\begin{enumerate}
    \item The client tool contacts a Bartender and send a `makeCollection' request with the LN \verb!/ourvo/common/docs!.
    \item The Bartender asks a Librarian to traverse this LN.
    \item The Librarian tries to traverse the Logical Name and it stops at the last possible point and returns the metadata of the last element. Because the user wants to put his collection in a new path but into an existing collection, we expect that only the \verb#/ourvo/common# part of the LN can be traversed. If the Librarian could traverse the whole LN that would mean that there is already an existing file or collection by that name. If LN does not exist but the the parent collection does, then the parent collection's metadata is returned to the Bartender. 
    \item The Bartender checks the access policies to decide if the user has permissions to put something into this collection. Then it asks the Librarian to create a new collection.
    \item The Librarian creates the collection, and returns its GUID. At this point this new collection has no real Logical Name, it only has a GUID, but it is not yet put into its parent collection.
    \item The Bartender now asks the Librarian to add this new entry into the parent collection, which means that the new GUID and the name \verb!docs! are added as a pair.
    \item The Librarian returns with a status message.
    \item Finally the Bartender tells the client tool if everything went OK or not.
\end{enumerate}

@@

A \emph{collection} is a list of files and other collections, which are in parent-children relationships forming a tree-hierarchy. Each entry has a unique name within a collection and a GUID which points to a corresponding file or collection, so a collection is basically a list of name-GUID pairs, which list is stored in the \emph{entries} section (see below). The collections could be in a closed state, which means that its contents should not be changed. If you close a collection then it cannot be opened again. However it cannot be guaranteed that the contents of a closed collection remains the same (e.g. if we include in our collection a file which we don't own, then the owner of the file could remove it nevertheless), but if a closed collection is changed, its state will be broken, and this state could never be changed again, which means you will always know if a closed collection is not intact anymore. This state is stored in the \emph{states} section, and its possible values are `\textbf{no}' if the collection is not closed, `\textbf{yes}' if the collection is closed and `\textbf{broken}' if the collection was closed, but something happened, and its contents have been changed.

@@

The internal representation of the access policies is based on access rules. A rule contains an identity and a list of actions. Currently the identity could be one of the following:
\begin{itemize}
    \item a Distinguish Name (DN) of a user, e.g. `\verb!/DC=eu/DC=KnowARC/O=NIIFI/CN=james!'
    \item a name of a Virtual Organization (VO) prefixed with `\verb!VOMS:!', e.g. `\verb!VOMS:knowarc.eu!'
    \item `\verb!ALL!' to specify that this rule will be applied to everyone
    \item `\verb!ANONYMOUS!' for a rule which will be applied to unauthenticated users (e.g. when non-secure HTTP is used and there is no DN)
\end{itemize}

@@

\note[JKN]{TODO: This paragraph is very unclear. All the 'could's makes it look like the TrustManager is just something planed in the far future}The TrustManager section itself could have some attributes: \textbf{FromFile} could have a filename, if we want to get the whole TrustManager configuration from a separate file; and \textbf{Force} indicates (with `yes' or `no') that we want to ensure that our service does not even get the request if the client of the incoming connection is not trusted.

@@

The A-Hash is a database for storing string tuples, and it provides conditional and atomic modification of them. The A-Hash can be used as a centralized service, or it could be deployed in a distributed way, using multiple nodes, where all the data is replicated on all the nodes.

The A-Hash stores \emph{objects}, where each object has an arbitrary string \emph{ID}, and contains any number of \emph{property}-\emph{value} pairs grouped in \emph{section}s, where \emph{property}, \emph{value} and \emph{section} are arbitrary strings. There could only be a single \emph{value} for a \emph{property} in a \emph{section}.

If you have an ID, you can get all property-value pairs of the corresponding object with the \emph{get} method, or you could specify only which sections or properties do you need. You can add or remove property-value pairs of an object or delete all occurrences of a property or create a new object with the \emph{change} method, and you can specify conditions, which means the change is only applied if the given conditions are met.

@@

    The \emph{IDs} is a list of string \emph{ID}'s, \emph{neededMetadataList} is a list of (\emph{section}, \emph{property}) pairs. If the \emph{neededMetadataList} is empty, then for each \emph{ID} it returns all the \emph{value}s for all the \emph{property} in all the \emph{section} in that given object. If there are sections and properties specified in \emph{neededMetadataList}, then only those values are returned.

@@

    The \emph{changeRequestList} is a list of \emph{changeRequest}s, where a \emph{changeRequest} is a tuple of (\emph{changeID}, \emph{ID}, \emph{changeType}, \emph{section}, \emph{property}, \emph{value}, \emph{conditionList}), where \emph{changeID} is an arbitrary ID which is used in the response to refer to this part of the request; \emph{ID} points to the object we want to change; \emph{changeType} can be `\textbf{set}' (to set the property within the section to value), `\textbf{unset}' (to remove the property from the section regardless of the value), `\textbf{delete}' (to remove the whole object); \emph{conditionList} is a list of \emph{condition}s, where a \emph{condition} is a tuple of (\emph{conditionID}, \emph{conditionType}, \emph{section}, \emph{property}, \emph{value}), where \emph{conditionType} could be `\textbf{is}' (which will be true if the property in the section is set to the value), `\textbf{isnot}' (which will be true if the property in the section is not set to the value), `\textbf{isset}' (which will be true if the property of the section is set, regardless of the value), `\textbf{unset}' (which will be true if the property of the section is not set at all).
    If all conditions are met, tries to apply changes to the objects, creates a new object if a previously non-existent ID is given.

    The response contains the \emph{changeResponseList} which is a list of (\emph{changeID}, \emph{success}, \emph{failedConditionID}) tuples. The \emph{success} of the change could be \textbf{set}, \textbf{unset}, \textbf{deleted}, \textbf{failed}, \textbf{condition not met} (in this case the ID of the failed condition is put into \emph{failedConditionID}), \textbf{invalid change type} or \textbf{unknown}.

@@

The A-Hash service has a modular architecture which means that you can decide if you want to deploy it a centralized or a distributed way by simply specifying different modules in the service's configuration. There are different low-level modules for storing the data on disk, e.g. serialized into `pickle' format. And there is a module which stores the data replicated on multiple nodes.

@@

\begin{description}
    \item[ReplicatedAHash] 
	is a subclass of the \verb!CentralizedAHash! business logic class and implements the extra logic needed for communication between A-Hash instances. The \verb!ReplicatedAHash! adds the method \verb!sendMessage! to the A-Hash interface and has a method \verb!processMessage! for processing messages from \verb!sendMessage!. These two methods are needed for the communication framework of the \verb!ReplicationManager!. The actual replication is managed by the \verb!ReplicationStore! and \verb!ReplicationManager!, meaning that the \verb!ReplicatedAHash!, with the exception of the above mentioned methods, is actually just a centralized A-Hash with a specific storage module.
    \item[ReplicationStore] 
	is a specialized replicated DB. It is a subclass of \verb!TransDBStore!\footnote{A module for storing data on disk transactionally using the Berkeley DB library.} with some additional features for handling replication. The \verb!ReplicationStore! periodically updates an A-Hash object containing information about the current master A-Hash and available clients. In addition it contains a \verb!ReplicationManager! object which governs the replication framework.
    \item[ReplicationManager] 
	handles all aspects of the Berkeley DB API and takes care of replicas, elections and replication messages.
	\begin{itemize}
	\item The replicated A-Hash consists of a set of replicas; one master and multiple clients. This means that the entire database is fully replicated on all replicas. When the master receives a write message, it will broadcast this message to all clients. To make sure that the db is consistent and that there will always be possible to elect a master, the master requires an acknowledgement message from a quorum of clients before continuing.
	\item In the replicated A-Hash there is one master and multiple clients. When the A-Hash starts up or if a master goes offline, we need to find a new master. This is solved by the clients holding an election, according to the Paxos algorithm. Simply told, all clients sends a vote to all other clients, and the eligible client that gets the most votes wins the election. An eligible client is a client which has received the latest updates. (See the Berkeley DB documentation for more details.) There are cases where some servers are more suitable to be master than others (better hardware, faster Ethernet, etc.). In such cases you can influence the election by setting a higher priority on the best servers. While an election is being held, it is not possible to write or read from the A-Hash, as consistency cannot be guaranteed without a master.
	\item As the replicas in the replicated A-Hash are services running within HED, the replication messages needs to be sent through the communication framework of HED. To be able to use this framework, the Berkeley DB needs a callback function for sending messages and the \verb!ReplicationManager! needs to call Berkeley DB to process incoming messages. The \verb!ReplicationMessage! takes a send function as one of the initialization arguments, and calls this function in the callback method. By providing the \verb!ReplicatedAHash.sendMessage! to the \verb!ReplicationManager! messages will be sent through the HED communication services just as in regular inter service communication. When \verb!ReplicatedAHash! receives a replication message from \verb!sendMessage! it will call \verb!ReplicationManager.processMessage! which in turn will call the Berkeley DB processMessage. This communication framework is both used by Berkeley DB and directly by the \verb!ReplicationManager! for, e.g., bootstrapping the A-Hash replicas and to communicate newly discovered clients.
\end{itemize}
\end{description}

@@

@@

\begin{description}
    \item[AHashClass] tells the A-Hash service the name of the business-logic A-Hash class. Currently there are two implementation, the \verb!CentralAHash! is a centralized version, and the \verb!ReplicatedAHash! is a replicated version.
    \item[StoreClass] specifies the name of the store class. This class is used to store the data on disk. Currently there are several versions (\verb!PickelStore!, \verb!CachedPickleStore!, \verb!StringStore!, \verb!ZODBStore!, \verb!TransDBStore!), however only the \verb!TransDBStore! is suitable for using in a replicated deployment, all the others are only for a centralized deployement. This variable is only configurable for the centralized deployement.
    \item[StoreCfg] could contain some parameters for the different store classes, and it almost always contains a \emph{DataDir} parameter which specifies a directory on the (local) filesystem where the A-Hash should save its files. This variable is only configurable for the centralized deployement.
\end{description}

For the replicated A-Hash there are some additional required configuration variables:

\begin{description}
    \item[LocalDir] specifies a directory on the local filesystem where this A-Hash replica should be stored. This directory must be unique for this replica, i.e., no two A-Hash replicas can share the same \verb!LocalDir!, and doing so will most likely leave both replicas broken beyond repair.
    \item[MyURL] is the URL to this A-Hash service and is used as an ID for this service in the replication group.
    \item[OtherURL] is the URL of another A-Hash in the replication group. More than one URL can be specified here.
\end{description}

Even though the default settings are useable, there are some optional variables available for the replicated A-Hash:
\begin{description}
    \item[Priority] is related to the master election. If a master dies, there will be an election to decide who will be the new master. Higher priority means a higher chance to be elected. A priority of 0 means that this replica can never be elected. Defaults to 10.
    \item[CheckPeriod] specifies in seconds how frequently the A-Hash should update the replica list. Defaults to 10 seconds.
    \item[CacheSize] configures the amount of the storage to be cached in memory. Defaults to 10MB
\end{description}

@@

Example of replicated A-Hash:
\begin{verbatim}
    <Service name="pythonservice" id="ahash1">
        <ClassName>storage.ahash.ahash.AHashService</ClassName>
        <AHashClass>storage.ahash.replicatedahash.ReplicatedAHash</AHashClass>
        <LocalDir>ahash_data1</LocalDir>
        <MyURL>http://localhost:60000/AHash1</MyURL>
        <OtherURL>http://localhost:60001/AHash2</OtherURL>
        <Priority>50</Priority>
        <ClientSSLConfig FromFile='clientsslconfig.xml'/>
    </Service>
\end{verbatim}

@@

\section{Librarians} % (fold)
\label{sec:librarian}

@@

The Librarian knows about the type of possible entries in the namespace of the ARC storage system: files, collections and mount points. It knows that these are organized in a tree-hiearchy with grouping them into collections, and it knows about Logical Names, and how to traverse a Logical Name to find the metadata of the file or collection (or mount point) which it refers to. However, the Librarian does not maintain the namespace of the ARC storage, e.g. it does not put new files into collections, or remove the link from its parent collection when a file is removed - this is the job of the Bartender. The only Librarian method which deals with Logical Names is the \emph{traverseLN} method, which traverses the Logical Name as long as possible, and returns the GUID and metadata of the last found entry in the path of the Logical Name. All the other methods work with GUID's, and do not care about the hierarchical namespace. In this regard the Librarian can be treated as a higher level application-specifc interface to the metadata store (the A-Hash), because while the A-Hash does not care about what it stores, the Librarian knowa exactly that it stores metadata about files, collections, etc. With this interface you can create new files, collections and mount points with the \emph{new} method, you can get or modify the metadata of any entry with the \emph{get} and the \emph{modifyMetadata} methods, and you can remove an entry with the \emph{remove} method.

Besides being an interface to the metadata store, the Librarian has an other importat functionality: it provides a way for Shepherds to register themselves, and to periodically send reports about the state of their stored files. A Shepherd send a \emph{report} message to a Librarian if e.g. a new file arrived, or an old file gets corrupted, thus the Librarian can keep the states of the replicas up-to-date. The Librarian monitors the registered Shepherds, and if one of them stops sending reports, the Librarian will assume, that it is offline, and modify the states of the all the replicas which are on the Shepherd's storage node. In order to do this, the Librarian stores for each Shepherd a list of files on the given storage nodes, this make it possible to know which files have replicas on a given storage nodes without checking all the files in the system. This list of files gets updated every time a Shepherd sends a report.

@@

    \item [get(getRequestList, neededMetadataList)] returns all or some of the metadata of the the requested GUID's.

    The \emph{getRequestList} is a list of string \emph{GUID}'s, \emph{neededMetadataList} is a list of (\emph{section}, \emph{property}) pairs. If the \emph{neededMetadataList} is empty, then for each \emph{GUID} it returns all the metadata \emph{value}s for all the \emph{property}s in all the \emph{section}s. If there are sections and properties specified in \emph{neededMetadataList}, then only those values are returned.

    The response contains a list of \emph{GUID, metadatList} pairs, where \emph{metadataList} is a list of (\emph{section}, \emph{property}, \emph{value}) tuples.

    \item[new(newRequestList)] creates a new entry in the Librarian with the given metadata, and returns the GUID of the new entry.

@@

    \emph{modifyMetadataRequestList} is a list of (\emph{changeID}, \emph{GUID}, \emph{changeType}, \emph{section}, \emph{property}, \emph{value}) tuples where \emph{changeType} can be `\textbf{set}' (set the property in the section to value), `\textbf{unset}' (remove the property-value pair from the section), `\textbf{add}' (set the property in the section to value only if it does not exist already) or `\textbf{setifvalue}=value' (only set the property in the section to value if it currently equals to the `value' in the \emph{changeType}).

    The response is a list of (\emph{changeID}, \emph{success}) where \emph{success} can be `\textbf{set}', `\textbf{unset}', `\textbf{condition failed}', `\textbf{failed}: reason'.

@@

    The response is a list of (\emph{requestID}, \emph{success}) pairs where \emph{success} could be `\textbf{removed}' or `\textbf{failed}: reason'.

@@

    The \emph{traverseRequestList} is a list of (\emph{requestID}, \emph{LN}) with the Logical Names to be traversed

>>>>

The different replicas of a file can be in different \note[JKN]{TODO: which states? a reference to definition of states maybe?}states. The replica can be valid and alive or just in the process of creation or it may be corrupt or a whole storage node may be offline. This state is always stored as metadata of the given replica. For each file there is a checksum calculated, and this checksum is used to detect if a replica gets corrupted. If a storage node (more precisely: the Shepherd service on the storage node) detects that a file is invalid, it reports this so the metadata will be in sync with the real state. And the storage nodes sends heartbeat messages periodically, and if a storage node goes offline, the missing heartbeat triggers the modification of metadata as well.

@@

\begin{enumerate}
    \item The client tool needs to know the URL of a Bartender. This can be preconfigured on the client side or, in future releases it may be acquired from an information system. When the client tool has the URL, it sends a `list' request which contains the Logical Name \verb!/ourvo/users/user!.
    \item The Bartender tries to get the metadata of the given LN by sending a `traverseLN' request to a Librarian.
    \item The Librarian service starts the traversing by asking an A-Hash service about the first part of the LN, which is the \verb!/! root collection. The A-Hash service only knows about GUID's and not about LN's, but the GUID of the root collection is well-known, so the A-Hash can return the metadata of the list of files and sub-collections in the root collection.
    \item Hopefully, the \verb!ourvo! collection can be found in the root collection, which means that the Librarian knows its GUID, and can ask for its metadata from the A-Hash.
    \item After the A-Hash returns the metadata of the \verb!/ourvo! collection, the Librarian finds the GUID of \verb!users! in it, then gets its metadata.
    \item The A-Hash returns the metadata of \verb!/ourvo/users! which contains the GUID of \verb!user!, so the Librarian can ask for its metadata.
    \item At last the A-Hash returns the metadata of \verb!/ourvo/users/user! to the Librarian, and the Librarian returns it to the Bartender. This metadata contains the list of entries within this collection, and it also contains the access policies for this collection.
    \item The Bartender first checks, based on the user's DN (or his VO membership) and the access policies of this collection, if the user has permission to get the contents of this collection or not. If the user is approved, then because the `list' request should return additional metadata about each entry in the collection, the Bartender sends a `get' message to a Librarian which requests the metadata of all the entries in this collection.
    \item The Librarian gets the metadata from the A-Hash and returns it to the Bartender.
    \item Now the Bartender has all the needed information and returns the proper response to the client tool which formats and prints the results nicely for the user.
\end{enumerate}

@@

\begin{enumerate}
    \item The client tool contacts a Bartender and send a `makeCollection' request with the LN \verb!/ourvo/common/docs!.
    \item The Bartender asks a Librarian to traverse this LN.
    \item The Librarian tries to traverse the Logical Name and it stops at the last possible point and returns the metadata of the last element. Because the user wants to put his collection in a new path but into an existing collection, we expect that only the \verb#/ourvo/common# part of the LN can be traversed. If the Librarian can traverse the whole LN it means that there is already an existing file or collection by that name. If LN does not exist but the the parent collection does, then the parent collection's metadata is returned to the Bartender. 
    \item The Bartender checks the access policies to decide if the user has permissions to put something into this collection. Then it asks the Librarian to create a new collection.
    \item The Librarian creates the collection, and returns its GUID. At this point this new collection has no real Logical Name, it only has a GUID, but it is not yet put into its parent collection.
    \item The Bartender now asks the Librarian to add this new entry into the parent collection, which means that the new GUID and the name \verb!docs! are added as a pair.
    \item The Librarian returns with a status message.
    \item Finally the Bartender tells the client tool if everything went OK or not.
\end{enumerate}

@@

A \emph{collection} is a list of files and other collections, which are in parent-children relationships forming a tree-hierarchy. Each entry has a unique name within a collection and a GUID which points to a corresponding file or collection, so a collection is basically a list of name-GUID pairs, which list is stored in the \emph{entries} section (see below). The collections can be in a closed state, which means that its contents should not be changed. If you close a collection then it cannot be opened again. However it cannot be guaranteed that the contents of a closed collection remains the same (e.g. if we include in our collection a file which we don't own, then the owner of the file can still remove it), but if a closed collection is changed, its state will be broken, and this state can never be changed again, which means you will always know if a closed collection is not intact anymore. This state is stored in the \emph{states} section, and its possible values are `\textbf{no}' if the collection is not closed, `\textbf{yes}' if the collection is closed and `\textbf{broken}' if the collection was closed, but something happened, and its contents have been changed.

@@

The internal representation of the access policies is based on access rules. A rule contains an identity and a list of actions. Currently the identity can be one of the following:
\begin{itemize}
    \item a Distinguish Name (DN) of a user, e.g. `\verb!/DC=eu/DC=KnowARC/O=NIIFI/CN=james!'
    \item a name of a Virtual Organization (VO) prefixed with `\verb!VOMS:!', e.g. `\verb!VOMS:knowarc.eu!'
    \item `\verb!ALL!' to specify that this rule will be applied to everyone
    \item `\verb!ANONYMOUS!' for a rule which will be applied to unauthenticated users (e.g. when non-secure HTTP is used and there is no DN)
\end{itemize}

@@

\note[JKN]{TODO: This paragraph is very unclear. All the 'could's makes it look like the TrustManager is just something planed in the far future. Are the attributes mandatory or optional?}The TrustManager section itself could have some attributes: \textbf{FromFile} could have a filename, if we want to get the whole TrustManager configuration from a separate file; and \textbf{Force} indicates (with `yes' or `no') that we want to ensure that our service does not even get the request if the client of the incoming connection is not trusted.

@@

The A-Hash is a database for storing string tuples, and it provides conditional and atomic modification of them. The A-Hash can be used as a centralized service, or it can be deployed in a distributed way, using multiple nodes, where all the data is replicated on all the nodes.

The A-Hash stores \emph{objects}, where each object has an arbitrary string \emph{ID}, and contains any number of \emph{property}-\emph{value} pairs grouped in \emph{section}s, where \emph{property}, \emph{value} and \emph{section} are arbitrary strings. It can only be one \emph{value} for a \emph{property} in a \emph{section}.

Given an ID, the \emph{get} method returns property-value pairs of the corresponding object. By specifying a section or property, the \emph{get} method returns the corresponding values. Using the \emph{change} method, property-value pairs can be added and removed from an object, all occurrences of a property can be deleted and new objects can be created. By specifying conditions, the change is only applied if the given conditions are met.

@@

    The \emph{IDs} is a list of string \emph{ID}'s, \emph{neededMetadataList} is a list of (\emph{section}, \emph{property}) pairs. If the \emph{neededMetadataList} is empty, then for each \emph{ID} it returns all the \emph{values} for all the \emph{properties} in all the \emph{section} in that given object. If there are sections and properties specified in \emph{neededMetadataList}, then only those values are returned.

@@

    The \emph{changeRequestList} is a list of \emph{changeRequest}s, where a \emph{changeRequest} is a tuple of (\emph{changeID}, \emph{ID}, \emph{changeType}, \emph{section}, \emph{property}, \emph{value}, \emph{conditionList}):
\begin{itemize}
\item \emph{changeID} is an arbitrary ID which is used in the response to refer to this part of the request
\item \emph{ID} points to the object we want to change
\item \emph{changeType} can be `\textbf{set}' to set the property within the section to value, `\textbf{unset}' to remove the property from the section regardless of the value, `\textbf{delete}' to remove the whole object
\item \emph{conditionList} is a list of \emph{condition}s, where a \emph{condition} is a tuple of (\emph{conditionID}, \emph{conditionType}, \emph{section}, \emph{property}, \emph{value}), where \emph{conditionType} can be `\textbf{is}' which will be true if the property in the section is set to the value, `\textbf{isnot}' which will be true if the property in the section is not set to the value, `\textbf{isset}' which will be true if the property of the section is set regardless of the value or `\textbf{unset}' which will be true if the property of the section is not set at all.
\end{itemize}
    If all conditions are met, the A-Hash tries to apply changes to the objects, and creates a new object if a previously non-existent ID is given.

    The response contains the \emph{changeResponseList} which is a list of (\emph{changeID}, \emph{success}, \emph{failedConditionID}) tuples. The \emph{success} of the change is either \textbf{set}, \textbf{unset}, \textbf{deleted}, \textbf{failed}, \textbf{condition not met} (in which case the ID of the failed condition is put into \emph{failedConditionID}), \textbf{invalid change type} or \textbf{unknown}.

@@

The A-Hash service has a modular architecture which means that it can be deployed in a centralized or a distributed way by simply specifying different modules in the service's configuration. There are different low-level modules for storing the metadata on disk, e.g. serialized into the `pickle' format. In addition there is a module which stores the metadata replicated on multiple nodes.

@@

\begin{description}
    \item[ReplicatedAHash] 
	is a subclass of the \verb!CentralizedAHash! business logic class and implements the extra logic needed for communication between A-Hash instances. The \verb!ReplicatedAHash! adds the method \verb!sendMessage! to the A-Hash interface and has a method \verb!processMessage! for processing messages from \verb!sendMessage!. These two methods are needed for the communication framework of the \verb!ReplicationManager!. The actual replication is managed by the \verb!ReplicationStore! and \verb!ReplicationManager!, meaning that the \verb!ReplicatedAHash!, with the exception of the above-mentioned methods, is actually just a centralized A-Hash with a specific storage module.
    \item[ReplicationStore] 
	is a specialized replicated database. It is a subclass of \verb!TransDBStore!\footnote{A module for storing data on disk transactionally using the Berkeley DB library.} with some additional features for handling replication. The \verb!ReplicationStore! periodically updates an A-Hash object which contains information about the current master A-Hash and available clients. In addition it contains a \verb!ReplicationManager! object which governs the replication framework.
    \item[ReplicationManager] 
	handles all aspects of the Berkeley DB API and takes care of replicas, elections of the master replica and replication messages.
	\begin{itemize}
	\item The replicated A-Hash consists of a set of replicas; one master and multiple clients. This means that the entire database is fully replicated on all replicas. When the master receives a write message, it will broadcast this message to all the clients. To make sure that the database is consistent and that it will always be possible to elect a master, the master requires an acknowledgement message from a quorum of clients before continuing.
	\item In the replicated A-Hash there is one master and multiple clients. When the A-Hash starts up or if a master goes offline, a new master must be elected. This is solved by the clients holding an election, according to the Paxos algorithm\footnote{http://research.microsoft.com/users/lamport/pubs/pubs.html\#lamport-paxos}. Simply told, all clients send a vote to all other clients, and the eligible client that gets the most votes wins the election. An eligible client is a client which has received the latest updates. See the Berkeley DB documentation for more details. There are cases where some servers are more suitable to be master than others (better hardware, faster Ethernet, etc.). In such cases the election can be influenced by setting a higher priority on the best servers. While an election is being held, it is not possible to write to or read from the A-Hash, as consistency cannot be guaranteed without a master.
	\item As the replicas in the replicated A-Hash are services running within HED, the replication messages need to be sent through the communication framework of HED. To be able to use this framework, the Berkeley DB needs a callback function for sending messages and the \verb!ReplicationManager! needs to call Berkeley DB to process incoming messages. The \verb!ReplicationMessage! takes a send function as one of the initialization arguments, and calls this function in the callback method. By providing the \verb!ReplicatedAHash.sendMessage! to the \verb!ReplicationManager! messages will be sent through the HED communication services just as in regular inter service communication. When \verb!ReplicatedAHash! receives a replication message from \verb!sendMessage! it will call \verb!ReplicationManager.processMessage! which in turn will call the Berkeley DB processMessage. This communication framework is both used by Berkeley DB and directly by the \verb!ReplicationManager!, e.g., for bootstrapping the A-Hash replicas and to communicate newly discovered clients.
\end{itemize}
\end{description}

@@

\subsection{Configuration} % (fold)
\label{sub:a_hash_configuration}

@@

\begin{description}
    \item[AHashClass] tells the A-Hash service the name of the business-logic A-Hash class. Currently there are two implementation, the \verb!CentralAHash! is a centralized version, and the \verb!ReplicatedAHash! is a replicated version.
    \item[StoreClass] specifies the name of the store class, used to store the data on disk. Currently there are several store classes \note[JKN]{TODO: describe the stores and how to choose among them}(\verb!PickelStore!, \verb!CachedPickleStore!, \verb!StringStore!, \verb!ZODBStore!, \verb!TransDBStore!), however only the \verb!TransDBStore! class is suitable for use in a replicated deployment. All the others are only for a centralized deployement. This variable is only configurable for the centralized deployement.
    \item[StoreCfg] contains parameters for the different store classes, and it almost always contains a \emph{DataDir} parameter which specifies a directory on the (local) filesystem where the A-Hash can save its files. This variable is only configurable for the centralized deployment.
\end{description}

For the replicated A-Hash there are some additional mandatory configuration variables:

\begin{description}
    \item[LocalDir] specifies a directory on the local filesystem where this A-Hash replica should be stored. This directory must be unique for this replica, i.e., no two A-Hash replicas can share the same \verb!LocalDir! since doing so will most likely leave both replicas broken beyond repair.
    \item[MyURL] is the URL to this A-Hash service and is used as an ID for this service in the replication group.
    \item[OtherURL] is the URL of another A-Hash in the replication group. More than one URL can be specified here.
\end{description}

Even though the default settings are useable, there are some optional variables available for the replicated A-Hash:
\begin{description}
    \item[Priority] is related to the master election. If a master dies, there will be an election to decide which client will be the new master. Higher priority means a higher chance to be elected. A priority of 0 means that this replica can never be elected. The defaults is 10.
    \item[CheckPeriod] specifies in seconds how frequently the A-Hash should update the replica list. The default is 10 seconds.
    \item[CacheSize] configures the amount of the storage to be cached in memory. The default is 10MB
\end{description}

Example of a centralized A-Hash configuration:

@@

Example of a replicated A-Hash configuration:
\begin{verbatim}
    <Service name="pythonservice" id="ahash1">
        <ClassName>storage.ahash.ahash.AHashService</ClassName>
        <AHashClass>storage.ahash.replicatedahash.ReplicatedAHash</AHashClass>
        <LocalDir>ahash_data1</LocalDir>
        <MyURL>http://localhost:60000/AHash1</MyURL>
        <OtherURL>http://localhost:60001/AHash2</OtherURL>
        <Priority>50</Priority>
        <ClientSSLConfig FromFile='clientsslconfig.xml'/>
    </Service>
\end{verbatim}

@@

\section{Librarian} % (fold)
\label{sec:librarian}

@@

The Librarian knows about the type of possible entries in the namespace of Chelonia: files, collections and mount points. It knows that these are organized in a tree-hiearchy with entries grouped into collections, and it knows about Logical Names, and how to traverse a Logical Name to find the metadata of the file or collection (or mount point) which it refers to. However, the Librarian does not maintain the namespace of Chelonia. It does not put new files into collections, nor remove the link from its parent collection when a file is removed - this is the job of the Bartender. The only Librarian method which deals with Logical Names is the \emph{traverseLN} method, which traverses the Logical Name as far as possible, and returns the GUID and metadata of the last found entry in the path of the Logical Name. All the other methods work with GUID's, and do not care about the hierarchical namespace. In this regard the Librarian can be treated as a higher level application-specifc interface to the metadata store (the A-Hash), because while the A-Hash does not care about what it stores, the Librarian knows exactly that it stores metadata about files, collections, etc. With this interface new files, collections and mount points can be created with the \emph{new} method, the metadata of any entry can be retrieved and modified with the \emph{get} and \emph{modifyMetadata} methods, and entries can be removed with the \emph{remove} method.

Besides being an interface to the metadata store, the Librarian has another important functionality: it provides a way for Shepherds to register themselves and to periodically send reports about the state of their stored files. A Shepherd sends a \emph{report} message to a Librarian if e.g. a new file arrived, or an old file gets corrupted, thus the Librarian can keep the states of the replicas up-to-date. The Librarian monitors the registered Shepherds, and if one of them stops sending reports, the Librarian will assume, that it is offline, and modify the states of all the replicas which are on the Shepherd's storage node. In order to do this, the Librarian stores for each Shepherd a list of files on the given storage nodes, this make it possible to know which files have replicas on a given storage nodes without checking all the files in the system. This list of files gets updated every time a Shepherd sends a report.

@@

    \item [get(getRequestList, neededMetadataList)] returns all or some of the metadata of the requested GUID's.

    The \emph{getRequestList} is a list of string \emph{GUID}'s, \emph{neededMetadataList} is a list of (\emph{section}, \emph{property}) pairs. If \emph{neededMetadataList} is empty, then for each \emph{GUID} all the metadata values for all the properties in all the sections are returned. If there are sections and properties specified in \emph{neededMetadataList}, then only those values are returned.

    The response contains a list of \emph{GUID, metadataList} pairs, where \emph{metadataList} is a list of (\emph{section}, \emph{property}, \emph{value}) tuples.

    \item[new(newRequestList)] creates a new entry in the Librarian with the given metadata, and returns the GUID's of the new entries.

@@

    \emph{modifyMetadataRequestList} is a list of (\emph{changeID}, \emph{GUID}, \emph{changeType}, \emph{section}, \emph{property}, \emph{value}) tuples where \emph{changeType} can be `\textbf{set}' to set the property in the section to value, `\textbf{unset}' to remove the property-value pair from the section, `\textbf{add}' to set the property in the section to \emph{value} only if it does not exist already or `\textbf{setifvalue}=value' to only set the property in the section to value if it currently equals to the `value' in the \emph{changeType}.

    The response is a list of (\emph{changeID}, \emph{success}) pairs where \emph{success} can be `\textbf{set}', `\textbf{unset}', `\textbf{condition failed}', `\textbf{failed}: reason'.

@@

    The response is a list of (\emph{requestID}, \emph{success}) pairs where \emph{success} is either `\textbf{removed}' or `\textbf{failed}: reason'.

@@

    The \emph{traverseRequestList} is a list of (\emph{requestID}, \emph{LN}) pairs with the Logical Names to be traversed

    The response is a list of (\emph{requestID}, \emph{traversedList}, \emph{wasComplete}, \emph{traversedLN}, \emph{GUID}, \emph{metadataList}, \emph{restLN}) tuples where:
    \begin{description}
        \item[traversedList] is a list of (\emph{LNpart}, \emph{GUID}) pairs, where \emph{LNpart} is a part of the \emph{LN}, \emph{GUID} is the GUID of the Librarian-entry referenced by that part of the \emph{LN}, the first element of this list is the shortest prefix of the \emph{LN}, the last element is the \emph{traversedLN} without its last part
        \item[wasComplete] indicates whether the full \emph{LN} was traversed
        \item[traversedLN] is the part of the \emph{LN} which was traversed, if \emph{wasComplete} is true, this should be the full \emph{LN}
        \item[GUID] is the \emph{GUID} of the \emph{traversedLN}
        \item[metadataList] is all the metadata of the of traversedLN in the form of (\emph{section}, \emph{property}, \emph{value}) tuples
        \item[restLN] is the remainders of the \emph{LN} which was not traversed for some reason, if \emph{wasComplete} is true, this should be an empty string
    \end{description}
